### 2024-01-01 thesis/chapters/02-analysis-B-features.tex
Fix grammar

\section{Signal preprocessing} \label{section:signal-preprocessing}
The vibration signals in a factory environment are inherently full of disturbances. Nearby equipment operation and handling of heavy objects in the surroundings can all contribute to the unwanted chaotic movement in otherwise mostly pure oscillatory motion. In addition, accelerometers suffer from systematic measurement errors in the form of thermal noise, zero-g offset as a result of slight miscalibration, and bias originating from a constant force of gravity. These unavoidable distortions are to some extent suppressible with digital filters. In the preprocessing stage, we consider detrending, noise reduction with adaptive filters, and time synchronous averaging to eliminate the external interference.

\subsection{Detrending}
The oscillatory motion should be centered around the zero level for further manipulation. The constant offset is eliminated simply by subtracting the overall mean from the signal. Moreover high pass DC blocker infinite impulse response (IIR) filter of 1st order can adjust to shifts of the average value over time (Equation~\ref{equ:iir-dc-blocker}). The transition band depends upon the choice of corner frequency $f_{3dB}$ (Fig.~\ref{fig:dc-blocker}).

@@

\section{Feature extraction} \label{section:feature-extraction}
After preprocessing raw numerical vectors are merely low-level descriptors of the underlying physical phenomena. At first, these incomprehensible sequences of numbers are reduced to summary attributes called features in the process of feature extraction or feature discovery. Features can be hand-crafted (as is our case), learned implicitly within the model representation, or explicitly from an optimization problem solution.

Predictive maintenance has ideal prerequisites for the application of feature engineering because the signal is usually pseudo-stationary, and the trend monitoring variables come out of extensive domain expertise in mechanics. The advantages of add-in extraction effort, as opposed to processing samples unmodified, are to gain better classification precision, reduce computational burden and storage capacity downstream with dimensionality reduction~\cite{johnson_feature_2019}.

It is important to note that the design of features is not a standalone step in the machine learning pipeline but it should be performed iteratively to improve the target model. Signal features are computed in the temporal and spectral domain~\cite{brito_fault_2021}.

\subsection{Temporal domain features}
The most widely found features in the literature are rudimentary statistical measures of the central moment: mean, variance, standard deviation, skewness, and kurtosis~(Tab.~\ref{tab:td-features}). Statistics can be calculated in any domain, but the mean value should not to be used in detrended data. The vibration severity metrics out of technical standards are also highly regarded. The characteristics of amplitude include root mean square, peak-to-peak distance, and maximum~\cite{mostafavi_novel_2021}.

The other significant time-domain attributes are derived as ratios of previous simpler ones. These ratios are crest factor, margin factor, impulse factor, and shape factor~(Tab.~\ref{tab:td-features})~\cite{nandi_condition_2019}. Many articles have been successful in bearing fault detection out of transients in impulsive signals with kurtosis, crest factor, and margin indicators \cite{brito_fault_2021}. It is also suggested that the shape factor can signify unbalance and misalignment faults~\cite{nandi_condition_2019}.

@@

In the spectral domain, we can obtain usual statistical properties of the distribution being spectral centroid, skew and kurtosis. Additionaly spectral roll-on and roll-off, fundamental frequency, entropy, negentropy, spectral flux, signal to noise ratio (noisiness), energy in frequency bands, and energy ratio are extracted (Tab.~\ref{tab:fd-features})~\cite{peeters_large_2004}.

In geometric terms, the spectral centroid represents the barycenter of the frequency magnitude plot. Spectral roll-off gives a notion about the spectral distribution because it identifies the frequency $f_c$ below which 95\% of the signal energy is contained. Complementary, the roll-on frequency is chosen so that 5\% of the signal energy is below this value. According to the definition, spectral flux is normalized cross-correlation between two successive amplitude spectra, a value of one means spectra are the most dissimilar.

@@

If a single principal frequency exists, it can be determined with maximum likelihood estimation. Such frequency would explain the signal spectrum the best~\cite{peeters_large_2004}.  The frequency spectrum is a discrete set of amplitudes where peaks have to be reliably identified to create representative attributes.

The essential peak-finding approaches are based either on magnitude or gradient. All found extrema are commonly filtered with the magnitude of prominences and the widths at half prominence. In the magnitude-based method, middle point $x_i$ is compared to neighboring two points and the peak is then: $x_{i-1} < x_i > x_{i+1}$. The gradient-based method evaluates the first derivative at the point which is equal to zero in case the point is a local maximum, local minimum, or inflection point~\cite{adikaram_non-parametric_2016}.

@@

Multiple harmonic series and the sidebands can be separated into a discrete set of frequency components, each with central frequency, uncertainty, and amplitude $C_i(v_i, \Delta v_i, A_i)$ by an exhaustive search algorithm. Harmonic family identification is a non-trivial problem because of the spectrum estimation errors. The criterion is proposed to select harmonic at the minimal distance from the true fundamental frequency multiple~(Equation~\ref{equ:harmonic-search}). Two series with the same fundamental frequency are merged and thought of as a modulation series~\cite{gerber_identification_2013}.

@@

\subsection{Time-frequency domain features}
The \textbf{Short-time Fourier transform} (STFT) splits the time-domain signal to an equal-length intervals. Individual chunks have 50\% overlap and are multiplied with weights of window function to balance scalloping loss and spectral leakage due to Fourier transform periodicity assumption. Traditionally, the Hann window is commonly used instead of rectangular window~\cite{ziaran_technicka_2013,noauthor_iso_2016_2}.

In the time-frequency domain, the same features can be derived as in the frequency domain, but in addition, the attributes are time-localized in this way. The STFT has a considerable flaw for implementation in a self-adaptable system and that is fixed resolution. The optimal window size has to be set beforehand or chosen after performing multiple transformations on chunks out of the range of lengths. Welch's methods averages multiple consecutive blocks to better estimate the spectrum. We have already researched the suitability of STFT for online detection of constant frequencies~\cite{hajek_iot_2022}.

The alternative to isolating weak impacts with high time resolution is a \textbf{Teager-Kaiser energy operator} (TKEO) (Equation~ \ref{equ:tkeo-operator}). It is a tool for envelope analysis to demodulate characteristic AM-FM signal present during bearing faults. Energy operator output can be utilized as a standalone feature attribute.

@@

Improved TKEO is necessary to prevent analysis from suffering from the noisy source. The key idea is to perform TKEO after signal decomposition into narrow-band components with different center frequencies. The extracted modes are reconstructed with weights assigned based on their correlations to the original signal~\cite{shi_application_2022}.

Time-frequency spectrum modification preserving localization of abrupt wide-band spikes at time $t_0$ and simultaneously reducing energy smear over the larger region is a goal of \textbf{Transient-extracting transform} (TET). Post-processing of STFT window $G(t,\omega)$ involves multiplication of the spectrum with the Transient extracting operator (TEO). This operator is expressed in the form of a Dirac delta function $\delta(t)$ (Equation~\ref{equ:tet-transform}).

@@

TET representation retains non-zero coefficients where the absolute value of the ratio between two STFTs $G^{tg}[n, k]\;/\;G[n, k]$ is less than half the sampling interval $T$. These two transforms use distinct windows $g[n]$ and $n \cdot g[n]$. Decomposition of signal with TET is proved to produce significantly larger kurtosis (around 38 in TET, 4 in other methods) and hence better discriminate the transient fault~\cite{yu_concentrated_2020}.

\subsection{Wavelet domain features}
The bands for lower frequencies should be longer in duration than for higher frequencies. The \textbf{Wavelet transform} (WT) possesses such multi-scale discrimination property effectively increasing resolution in time-frequency plain. Wavelet basis functions are constructed for that purpose (Equation~\ref{equ:mother-wavelet}). There are several wavelet families, for example, Haar, Daubechies, Coiflets, Symlets, Morlet, and Meyer~\cite{nandi_condition_2019}.

@@

\textbf{Continuous Wavelet transform} (CWT)~(Equation~\ref{equ:cwt-transform}) is performed by scaling and translating the mother wavelet $\psi$ picked out of the appropriate family~\cite{nandi_condition_2019}. The scale factor is denoted with~$s$ and time position with~$\tau$. The choice of wavelet type is data-driven because distinct wavelet shapes have an impact on the response and ultimately contribute to filter length. The decision lies between recognition abilities for impulse-like signals or the inclusion of wider surrounding space.

@@

The CWT is computationally intensive when a highly detailed scale resolution is required because each wavelet scale convolves with the entire signal. The fCWT algorithm allows 100 times higher spectral resolution then previous implementation at the same speed. It increases performance 122 times compared to Wavelib and 34 times in comparison with PyWavelets~\cite{arts_fast_2022}.

The fast CWT algorithm reaches compelling improvement by applying Parseval's theorem to the wavelet transform formula that removes the dependence on the time offset parameter~\cite{arts_fast_2022}. The convolution takes place with the mother wavelet in the Fourier base. Then, inverse FFT produces the coefficients for individual scales.

@@

The spectrograms to illustrate the difference in the ability of Fourier transform, continuous Wavelet transform, and their modifications to pick up underlying patterns in bearing faults are shown in the Fig.~\ref{fig:transforms}.

@@

The dyadic filter bank is another signal decomposition technique that generates subbands on multiple granularity levels. The practical realization of the multi-scale description is \textbf{Discrete Wavelet Transform} (DWT). The DWT behaves as a quadrature mirror filter and splits waveform using a wavelet filter to detail coefficients (D1) and approximation coefficients (A1) (Fig.~\ref{fig:dwt-filter-bank})~\cite{nandi_condition_2019}. Low-pass filter $h(k)$ creates approximation coefficients further decomposed in the successive levels. Detail coefficients represent the result of the high-pass filter $g(k) = (-1)^k h(1 - k)$ after decimation by the factor of 2.

@@

In washing machine status classification, the discrete wavelet transform with Daubechies wavelet (db4) and fifth-level decomposition provided features combined from approximation (cA5) and detail coefficients (cD1, \dots, cD5). Washing machines belonged to three categories: no fault, electric motor clamping screws problem, and a loose or broken counterweight. Extracted measures were sample mean and sample variances over autocorrelation functions of coefficients (AcD$n$) and smoothed coefficients cD1, cD2 by moving average filter~\cite{goumas_classification_2002}.

@@

Each wavelet packet coefficient $w_{j,n,k}$ captures subband frequency content around time instant $2^j k$ (Equation~\ref{equ:wavelet-packet-coefficient})~\cite{yen_wavelet_2000}. This measure is an inner product of the source and scaled wavelet packet function. The aforementioned feature extraction established in DWT can be applied, for example calculation of the wavelet packet node energy.

@@

The wavelet packet energy ratio for weak feature extraction has been incorporated into the method of multiple frequency bands demodulation (MFBD). The highest $n$ energy coefficients are selected out of 8 narrow frequency bands to subsequently affect the principal components. Demodulation principle for the frequency bandwidth prescribes $n$ to satisfy condition: $1\;/\;2^n > f_{\mathrm{modulation}}\;/\;f_s$. The first few eigenvectors explaining together more than 80\% of energy are retained to reconstruct the signal with Fourier transform and retain the weak fault components~\cite{song_mfbd_2021}.

@@

The normalized frequency axis in range $\omega \in [0, \pi]$ is divided by split points $\omega_0, \dots, \omega_N$ where $\omega_n = f_n \cdot 2 \pi\;/\;f_s$ (Fig.~\ref{fig:ewt-spectrum-segmentation}). Each segment is bounded between $[\omega_{n-1}, \omega_n]$ with transition phase of width $2\tau_n$ and polynomial transition function $\beta(x)$. A tight frame set of empirical wavelets is built by setting the transition phase proportional to the band boundary: $\tau_n = \lambda \omega_n$, and proportional constant must obey constrain: $\lambda = \min\left(\frac{\omega_{n+1} - \omega_n}{\omega_{n+1} + \omega_n}\right)$. The $N$ boundaries defining different portions of the Fourier spectrum are placed at the center between two consecutive local maxima~\cite{gilles_empirical_2013}.

@@

The drawback of EWT is improper segmentation in noisy and non-stationary signals producing too many uninformative partitions of the spectrum. The solutions we discuss fall into two groups that apply improvements before or after the wavelet transformation step.

A combination of \textbf{Maximum Correlated Kurtosis Deconvolution} and improved EWT (MKCD-EWT) favors periodic impacts by dynamically constructing an optimal FIR filter that maximizes the correlation kurtosis of the signal. The envelope curve smooths of the amplitude spectrum by linear interpolation (Fig.~\ref{fig:mkcd-ewt-segmentation}). The threshold accounting for desired SNR modifies the envelope when more than $n$ components are discovered ($A_h$ is maximal amplitude and $A_l$ is minimal amplitude): $\lambda = A_l + \frac{C}{\mathrm{SNR}}(A_h - A_l)$. \emph{Squared envelope spectrum} and \emph{Teager energy operator spectrum} are computed as fault features for the IMF with highest kurtosis~\cite{li_fault_2019}.

The segmentation boundaries detection can be accomplished also on the spectral envelope consisting of \textbf{Piecewise Cubic Hermite Interpolating Polynomial} (PCHIP-EWT) instead of scanning the Fourier spectrum directly (Fig.~\ref{fig:pchip-ewt-segmentation}). Divisions of the upper cut-off frequency are kept if local power indicates a subband with useful information ($p_i(f) \geq \lambda$)  as in the case of abrupt level change. Local power is the ratio of $n$th local envelope maximum and difference of indexes of its adjacent local minima:  $p_i(f) = K_{max(i)(f)} \;/\;\left(f_{k_{min(i+1)}} - f_{k_{min(i)}}\right)$. The best $\lambda$ is set experimentally~\cite{zhuang_improved_2020}.

@@

The inspiration for filtering with adaptive basis comes from \textbf{Empirical mode decomposition} (EMD). This procedure locates local maxima and minima in the time waveform that are interpolated with cubic splines to form signal envelopes. The average of the upper and lower envelope is repeatedly subtracted from the residuals recovering the higher order and lower frequency IMFs~\cite{wang_computational_2014}.

@@

Wavelet domain features and associated spectral segmenation appears to be a powerful tool in conserving data bandwidth. For the use in white-box machinery diagnosis these trend indicator turn out to be difficult to comunicate and comprehend. Therefore we use more standardized metrics in the further study.

\section{Feature transformations} \label{section:feature-selection}
Numerical features from the feature extraction phase have non-normal distributions and span wide ranges. Broad differences among features skew the spread in the particular axis. Inevitably it can degrade the discernment of fault diagnostics models that map input onto smooth function as regression does~\cite{zheng_feature_2018}. The feature scaling, power transform, and principal component analysis modify attribute values to gain more meaningful predictors, but one must be cautious in model interpretation.

\subsection{Feature normalization}
Vibrations are measured with an accelerometer in the form of 3D vectors where each axis has its component. Models have to be resilient to any slight inclination or sensor orientation. To satisfy this prerequisite, feature $f$ extracted from all three dimensions is first composed into a single vector and Euclidian norm is computed~\cite{kamminga_robust_2018}.

@@

Feature normalization most often takes two forms. \textbf{Min-max scaling} changes original range of values into interval $[0, 1]$ (Equation~\ref{equ:min-max-scaler}). \textbf{Standardization} (Equation~\ref{equ:standardization-scaler}) constrains the mean of the variable to 0 with a variance of 1~\cite{zheng_feature_2018}.

@@

\subsection{Principal components analysis}
Highly correlated features generated from a relatively small original column space are redundant as they do not provide any additional information for diagnostics. \textbf{Principal Component Analysis} (PCA) solves this problem by projecting potentially linearly dependant features into a new feature space where the incoming information is preserved in a smaller number of features~\cite{zheng_feature_2018}. The threshold of how many principal components are picked depends on the amount of explained variance and desired quantity of data reduction.

@@

\section{Feature selection}
Features do not contribute to predictive power of the model with an even share. A certain subset can reach better results than others. Choosing optimal feature subset is NP-hard combinatorial problem.

@@

The general steps in selecting the adequate predictors is as follows~\cite{nandi_condition_2019}:
\begin{enumerate}
    \itemsep0pt
    \item \textbf{Subset generation} - sets of features are generated in different search directions and with various strategies. Attributes are either appended to an empty set or pruned away from a universal set, sequentially or randomly.
    \item \textbf{Subset evaluation} - comparison of subset quality is assessed with relevance measure some of which are discussed below.
    \item \textbf{Stopping criteria} - search is exhausted when the specified number of features has been found, subset metrics cannot be improved further, or satisfactory model performance is achieved. Subset generation and evaluation can be performed multiple times until the stopping criteria are met.
    \item \textbf{Validation} - resulting subset is tested for the specific model on synthetic and real-world datasets against well-known results. 
\end{enumerate}

@@

The frequently used scores upon which the feature relevance is ordered are~\cite{nandi_condition_2019}: 
\begin{itemize}
\item \textbf{Variance threshold} - removes low-variance features below set threshold.
\item \textbf{Correlation coefficient} - expresses the linear relationship between two variables. The codependent variables are of three sorts: quantitative, ordinal, and nominal. The choice of coefficient calculation is determined by the type of variables under consideration shown in Table \ref{tab:corr-coef}.

\emph{Pearson correlation} coeficient expresses similarity between two quantitative features $f_i$ and $f_j$. In the classification setting the correlation of feature $f$ makes sense only with dichotomous target class label $c$ using \emph{point biserial coeficient}. Mathematically, this is equivalent to the Pearson coefficient. Rank correspondence is quantified with either Spearman rho or Kendall's Tau~\cite{calkins_more_2005}.

@@

Attributes are ranked in descending order according to the absolute value of their correlation coefficient. We seek the highest correlation to the class label.
\myequations{Pearson correlation coefficient}
\begin{ceqn}\begin{align}
r(i) = \frac{\mathrm{cov}(f, c)}{\sqrt{\mathrm{var}(f) \cdot \mathrm{var}(c)}}
\end{align}\end{ceqn}

\item \textbf{Fisher score} - measures the difference among the means of the classes. It is interchangeable with ANOVA F-value, but it is evaluated for each feature $X^j$ separately. Ideally, the features in the subset have large distances between samples of various classes in $C$ and distances within a class are the smallest possible. In the formula (\ref{equ:fisher-score}), $n_j$ is the sample size of $j$th feature, $\mu^j$ is its sample mean, and $\mu$ is the overall mean.

>>>>

\section{Signal preprocessing} \label{section:signal-preprocessing}
The vibration signals in a factory environment are inherently full of disturbances. Nearby equipment operation and handling of heavy objects in the surroundings can all contribute to the unwanted chaotic movement in otherwise mostly pure oscillatory motion. In addition, accelerometers suffer from systematic measurement errors in the form of thermal noise, zero-g offset as a result of slight miscalibration, and bias originating from a constant force of gravity. These unavoidable distortions are to some extent suppressable by digital filters. In the preprocessing stage, we consider trend removal, noise reduction with adaptive filters, and time synchronous averaging to eliminate external interference.

\subsection{Detrending}
The oscillatory motion should be centered around the zero level for further manipulation. The constant offset is eliminated simply by subtracting the overall mean from the signal. Moreover, the high pass DC blocker infinite impulse response (IIR) filter of 1st order can adjust to shifts of the average value over time (Equation~\ref{equ:iir-dc-blocker}). The transition band depends upon the choice of corner frequency $f_{3dB}$ (Fig.~\ref{fig:dc-blocker}).

@@

\section{Feature extraction} \label{section:feature-extraction}
After preprocessing, raw numerical vectors are merely low-level descriptors of the underlying physical phenomena. At first, these incomprehensible sequences of numbers are reduced to summary attributes called features in the process of feature extraction or feature discovery. Features can be hand-crafted (as is our case), learned implicitly within the model representation, or explicitly from an optimization problem solution.

Predictive maintenance has ideal prerequisites for the application of feature engineering because the signal is usually pseudo-stationary, and the trend monitoring variables come out of extensive domain expertise in mechanics. The advantages of the add-in extraction effort, as opposed to processing samples unmodified, are to gain better classification precision, reduce computational burden and storage capacity downstream with dimensionality reduction~\cite{johnson_feature_2019}.

It is important to note that the design of features is not a standalone step in the machine learning pipeline, but it should be done iteratively to improve the target model. Signal features are computed in the temporal and spectral domain~\cite{brito_fault_2021}.

\subsection{Temporal domain features}
The most widely found features in the literature are rudimentary statistical measures of the central moment: mean, variance, standard deviation, skewness, and kurtosis~(Tab.~\ref{tab:td-features}). Statistics can be calculated in any domain, but the mean value should not be used in detrended data. The vibration severity metrics out of technical standards are also highly regarded. The characteristics of amplitude include root-mean-square, peak-to-peak distance, and maximum~\cite{mostafavi_novel_2021}.

The other significant time-domain attributes are derived as ratios of previous simpler ones. These ratios are crest factor, margin factor, impulse factor, and shape factor~(Tab.~\ref{tab:td-features})~\cite{nandi_condition_2019}. Many articles have been successful in fault detection of bearings out of transients in impulsive signals with kurtosis, crest factor, and margin indicators \cite{brito_fault_2021}. It is also suggested that the shape factor can signify unbalance and misalignment faults~\cite{nandi_condition_2019}.

@@

In the spectral domain, we can obtain the usual statistical properties of the distribution being spectral centroid, skew and kurtosis. Additionally, spectral roll-on and roll-off, fundamental frequency, entropy, negentropy, spectral flux, signal-to-noise ratio (noisiness), energy in frequency bands, and energy ratio are extracted (Tab.~\ref{tab:fd-features})~\cite{peeters_large_2004}.

In geometric terms, the spectral centroid represents the barycenter of the frequency magnitude plot. Spectral roll-off gives a notion about the spectral distribution because it identifies the frequency $f_c$ below which 95\% of the signal energy is contained. Complementary, the roll-on frequency is chosen so that 5\% of the signal energy is below this value. According to the definition, spectral flux is normalized cross-correlation between two successive amplitude spectra a value of one means the spectra are the most dissimilar.

@@

If a single principal frequency exists, it can be determined by maximum likelihood estimation. Such frequency would explain the signal spectrum the best~\cite{peeters_large_2004}. The frequency spectrum is a discrete set of amplitudes where peaks have to be reliably identified to create representative attributes.

The essential peak-finding approaches are based either on magnitude or gradient. All found extrema are commonly filtered with the magnitude of prominences and the width at half prominence. In the magnitude-based method, the middle point $x_i$ is compared to two neighboring points and the peak is then: $x_{i-1} < x_i > x_{i+1}$. The gradient-based method evaluates the first derivative at the point which is equal to zero if the point is a local maximum, local minimum, or inflection point~\cite{adikaram_non-parametric_2016}.

@@

Multiple harmonic series and the sidebands can be separated into a discrete set of frequency components, each with a central frequency, uncertainty, and amplitude $C_i(v_i, \Delta v_i, A_i)$ by an exhaustive search algorithm. Harmonic family identification is a non-trivial problem because of spectrum estimation errors. The criterion is proposed to select harmonics at the minimal distance from the true fundamental frequency multiple~(Equation~\ref{equ:harmonic-search}). Two series with the same fundamental frequency are merged and thought of as a modulation series~\cite{gerber_identification_2013}.

@@

\subsection{Time-frequency domain features}
The \textbf{Short-time Fourier transform} (STFT) splits the time-domain signal into equal-length intervals. Individual chunks have a 50\% overlap and are multiplied with weights of window function to balance scalloping loss and spectral leakage due to Fourier transform periodicity assumption. Traditionally, the Hann window is commonly used instead of the rectangular window~\cite{ziaran_technicka_2013,noauthor_iso_2016_2}.

In the time-frequency domain, the same features can be derived as in the frequency domain, but in addition, the attributes are time-localized in this way. The STFT has a considerable flaw for implementation in a self-adaptable system and that is fixed resolution. The optimal window size has to be set beforehand or chosen after performing multiple transformations on chunks out of the range of lengths. Welch's method averages multiple consecutive blocks to better estimate the spectrum. We have already researched the suitability of STFT for online detection of constant frequencies~\cite{hajek_iot_2022}.

The alternative to isolating weak impacts with high time resolution is a \textbf{Teager-Kaiser energy operator} (TKEO) (Equation~ \ref{equ:tkeo-operator}). It is a tool for envelope analysis to demodulate characteristic AM-FM signals present during bearing faults. Energy operator output can be utilized as a standalone feature attribute.

@@

Improved TKEO is necessary to prevent analysis from suffering from the noisy source. The key idea is to perform TKEO after signal decomposition into narrowband components with different center frequencies. The extracted modes are reconstructed with weights assigned based on their correlations to the original signal~\cite{shi_application_2022}.

Time-frequency spectrum modification preserving localization of abrupt wideband spikes at time $t_0$ and simultaneously reducing energy smear over the larger region is a goal of \textbf{Transient-extracting transform} (TET). Post-processing of STFT window $G(t,\omega)$ involves multiplication of the spectrum with the Transient extracting operator (TEO). This operator is expressed in the form of a Dirac delta function $\delta(t)$ (Equation~\ref{equ:tet-transform}).

@@

TET representation retains non-zero coefficients where the absolute value of the ratio between two STFTs, $G^{tg}[n, k]\;/\;G[n, k]$ is less than half the sampling interval $T$. These two transforms use distinct windows $g[n]$ and $n \cdot g[n]$. Decomposition of a signal with TET is proved to produce significantly larger kurtosis (around 38 in TET, 4 in other methods) and hence better discriminate the transient fault~\cite{yu_concentrated_2020}.

\subsection{Wavelet domain features}
The bands for lower frequencies should be longer in time than for higher frequencies. The \textbf{Wavelet transform} (WT) possesses such a multiscale discrimination property, effectively increasing resolution in time-frequency plain. Wavelet basis functions are constructed for that purpose (Equation~\ref{equ:mother-wavelet}). There are several wavelet families. For example, Haar, Daubechies, Coiflets, Symlets, Morlet, and Meyer~\cite{nandi_condition_2019}.

@@

\textbf{Continuous Wavelet transform} (CWT)~(Equation~\ref{equ:cwt-transform}) is done by scaling and translating the mother wavelet $\psi$ picked out of the appropriate family~\cite{nandi_condition_2019}. The scale factor is denoted with~$s$ and time position with~$\tau$. The choice of wavelet type is data-driven because distinct wavelet shapes have an impact on the response and ultimately contribute to filter length. The decision lies between recognition abilities for impulse-like signals or the inclusion of wider surrounding space.

@@

The CWT is computationally intensive when a highly detailed scale resolution is required because each wavelet scale convolves with the entire signal. The fCWT algorithm allows 100 times higher spectral resolution than previous implementations at the same speed. It increases performance 122 times compared to Wavelib and 34 times in comparison with PyWavelets~\cite{arts_fast_2022}.

The fast CWT algorithm reaches compelling improvement by applying Parseval's theorem to the wavelet transform formula that removes the dependence on the time offset parameter~\cite{arts_fast_2022}. The convolution takes place with the mother wavelet in the Fourier base. Then, the inverse FFT produces the coefficients for individual scales.

@@

The spectrograms to illustrate the difference in the ability of the Fourier transform, the continuous Wavelet transform, and their modifications to pick up underlying patterns in bearing faults are shown in Fig.~\ref{fig:transforms}.

@@

The dyadic filter bank is another signal decomposition technique that generates subbands at multiple granularity levels. The practical realization of the multiscale description is \textbf{Discrete Wavelet Transform} (DWT). The DWT behaves as a quadrature mirror filter and splits waveforms using a wavelet filter to detail coefficients (D1) and approximation coefficients (A1) (Fig.~\ref{fig:dwt-filter-bank})~\cite{nandi_condition_2019}. The low-pass filter $h(k)$ creates approximation coefficients further decomposed at the successive levels. Detail coefficients represent the result of the high-pass filter $g(k) = (-1)^k h(1 - k)$ after decimation by a factor of 2.

@@

In washing machine status classification, the discrete wavelet transform with Daubechies wavelet (db4) and fifth-level decomposition provide features with a combination of approximation (cA5) and detail coefficients (cD1, \dots, cD5). Washing machines belong to three categories: no fault, electric motor clamping screws problem, and a loose or broken counterweight. Extracted measures were sample mean and sample variances over autocorrelation functions of coefficients (AcD$n$) and smoothed coefficients cD1, cD2 by moving average filter~\cite{goumas_classification_2002}.

@@

Each wavelet packet coefficient $w_{j,n,k}$ captures subband frequency content around time instant $2^j k$ (Equation~\ref{equ:wavelet-packet-coefficient})~\cite{yen_wavelet_2000}. This measure is an inner product of the source and scaled wavelet packet function. The aforementioned feature extraction established in DWT can be applied. For example calculation of the wavelet packet node energy.

@@

The wavelet packet energy ratio for weak feature extraction has been incorporated into the method of multiple frequency band demodulation (MFBD). The highest $n$ energy coefficients are selected out of 8 narrow frequency bands to subsequently affect the principal components. Demodulation principle for the frequency bandwidth prescribes $n$ to satisfy condition: $1\;/\;2^n > f_{\mathrm{modulation}}\;/\;f_s$. The first few eigenvectors that explain together more than 80\% of energy are retained to reconstruct the signal with a Fourier transform and retain the weak fault components~\cite{song_mfbd_2021}.

@@

The normalized frequency axis in range $\omega \in [0, \pi]$ is divided by split points $\omega_0, \dots, \omega_N$ where $\omega_n = f_n \cdot 2 \pi\;/\;f_s$ (Fig.~\ref{fig:ewt-spectrum-segmentation}). Each segment is bounded between $[\omega_{n-1}, \omega_n]$ with a transition phase of width $2\tau_n$ and polynomial transition function $\beta(x)$. A tight frame set of empirical wavelets is built by setting the transition phase proportional to the band boundary: $\tau_n = \lambda \omega_n$, and proportional constant must obey constrain: $\lambda = \min\left(\frac{\omega_{n+1} - \omega_n}{\omega_{n+1} + \omega_n}\right)$. The $N$ boundaries defining different portions of the Fourier spectrum are placed at the center between two consecutive local maxima~\cite{gilles_empirical_2013}.

@@

The drawback of EWT is improper segmentation in noisy and non-stationary signals, producing too many uninformative partitions of the spectrum. The solutions we discuss fall into two groups that apply improvements before or after the wavelet transformation step.

A combination of \textbf{Maximum Correlated Kurtosis Deconvolution} and improved EWT (MKCD-EWT) favors periodic impacts by dynamically constructing an optimal FIR filter that maximizes the correlation kurtosis of the signal. The envelope curve smooths the amplitude spectrum by linear interpolation (Fig.~\ref{fig:mkcd-ewt-segmentation}). The threshold accounting for desired SNR modifies the envelope when more than $n$ components are discovered ($A_h$ is maximal amplitude and $A_l$ is minimal amplitude): $\lambda = A_l + \frac{C}{\mathrm{SNR}}(A_h - A_l)$. \emph{Squared envelope spectrum} and \emph{Teager energy operator spectrum} are computed as fault features for the IMF with the highest kurtosis~\cite{li_fault_2019}.

The segmentation boundary detection can also be accomplished on the spectral envelope consisting of \textbf{Piecewise Cubic Hermite Interpolating Polynomial} (PCHIP-EWT) instead of scanning the Fourier spectrum directly (Fig.~\ref{fig:pchip-ewt-segmentation}). Divisions of the upper cut-off frequency are kept if local power indicates a subband with useful information ($p_i(f) \geq \lambda$)  as in the case of an abrupt level change. Local power is the ratio of $n$th local envelope maximum and the difference between indexes of its adjacent local minima:  $p_i(f) = K_{max(i)(f)} \;/\;\left(f_{k_{min(i+1)}} - f_{k_{min(i)}}\right)$. The best $\lambda$ is set experimentally~\cite{zhuang_improved_2020}.

@@

The inspiration for filtering on an adaptive basis comes from \textbf{Empirical mode decomposition} (EMD). This procedure locates local maxima and minima in the time waveform that are interpolated with cubic splines to form signal envelopes. The average of the upper and lower envelope is repeatedly subtracted from the residuals recovering the higher order and lower frequency IMFs~\cite{wang_computational_2014}.

@@

Wavelet domain features and associated spectral segmentation appear to be a powerful tool in conserving data bandwidth. These variables turn out to be difficult to communicate and comprehend. Therefore, we will use more standardized metrics in further studies.

\section{Feature transformations} \label{section:feature-selection}
Numerical features from the feature extraction phase have non-normal distributions and span wide ranges. Broad differences among features skew the spread on a particular axis. Inevitably, it can degrade the discernment of fault diagnostics models that map input onto a smooth function as regression does~\cite{zheng_feature_2018}. The feature scaling, power transform, and principal component analysis modify attribute values to gain more meaningful predictors, but one must be cautious in model interpretation.

\subsection{Feature normalization}
Vibrations are measured with an accelerometer in the form of 3D vectors where each axis has its own component. Models have to be resilient to any slight inclination or sensor orientation. To satisfy this prerequisite, feature $f$ extracted from all three dimensions is first composed into a single vector and the Euclidian norm is computed~\cite{kamminga_robust_2018}.

@@

Feature normalization most often takes two forms. \textbf{Min-max scaling} changes the original range of values into an interval $[0, 1]$ (Equation~\ref{equ:min-max-scaler}). \textbf{Standardization} (Equation~\ref{equ:standardization-scaler}) constrains the mean of the variable to 0 with a variance of 1~\cite{zheng_feature_2018}.

@@

\subsection{Principal components analysis}
Highly correlated features generated from a relatively small original column space are redundant as they do not provide any additional information for diagnostics. \textbf{Principal Component Analysis} (PCA) solves this problem by projecting potentially linearly dependent features into a new feature space where the incoming information is preserved in a smaller number of features~\cite{zheng_feature_2018}. The threshold of how many principal components are picked depends on the amount of explained variance and desired quantity of data reduction.

@@

\section{Feature selection}
Features do not contribute to the predictive power of the model with an even share. A certain subset can achieve better results than others. Choosing the optimal feature subset is the NP-hard combinatorial problem.

@@

The general steps in selecting the appropriate predictors are as follows~\cite{nandi_condition_2019}:

\begin{enumerate}
    \itemsep0pt
    \item \textbf{Subset generation} - sets of features are generated in different search directions and with various strategies. Attributes are either appended to an empty set or pruned away from a universal set, sequentially or randomly.

    \item \textbf{Subset evaluation} - comparison of subset quality is assessed with relevance measures, some of which are discussed below.

    \item \textbf{Stopping criteria} - search is exhausted when the specified number of features has been found, subset metrics cannot be improved further, or satisfactory model performance is achieved. Subset generation and evaluation can be performed multiple times until the stopping criteria are met.

    \item \textbf{Validation} - the resulting subset is tested for the specific model on synthetic and real-world datasets against well-known results. 
\end{enumerate}

@@

The frequently used scores upon which the feature relevance is ordered are~\cite{nandi_condition_2019}:

\begin{itemize}
\item \textbf{Variance threshold} - removes low-variance features below the set threshold.

\item \textbf{Correlation coefficient} - expresses the linear relationship between two variables. The codependent variables are of three sorts: quantitative, ordinal, and nominal. The choice of coefficient calculation is determined by the type of variables under consideration shown in Table \ref{tab:corr-coef}.

\emph{Pearson correlation} coeficient expresses similarity between two quantitative features: $f_i$ and $f_j$. In the classification setting, the correlation of feature $f$ makes sense only with dichotomous target class label $c$ using \emph{point biserial coeficient}. Mathematically, this is equivalent to the Pearson coefficient. Rank correspondence is quantified with either Spearman rho or Kendall's Tau~\cite{calkins_more_2005}.

@@

Attributes are ranked in descending order according to the absolute value of their correlation coefficient. We seek the highest correlation to the class label.

\myequations{Pearson correlation coefficient}
\begin{ceqn}\begin{align}
r(i) = \frac{\mathrm{cov}(f, c)}{\sqrt{\mathrm{var}(f) \cdot \mathrm{var}(c)}}
\end{align}\end{ceqn}

\item \textbf{Fisher score} - measures the difference between the means of the classes. It is interchangeable with ANOVA F-value, but it is evaluated for each feature $X^j$ separately. Ideally, the features in the subset have large distances between samples of various classes in $C$ and distances within a class are the smallest possible. In the formula (\ref{equ:fisher-score}), $n_j$ is the sample size of $j$th feature, $\mu^j$ is its sample mean, and $\mu$ is the overall mean.
